### GTEx polygenic risk score pipeline for approved MRC-IEU users  

### Installing Package


```{r,}
# install devtools
install.packages('devtools')
library(devtools)
```

```{r,}
# install package from github
install_github("mbyvcm/MRCIEUGTEx", quiet = T)
library(MRCIEUGTEx)
```
### Adding directories to GTEx data on RDSF

These can be added manually if you know where GTEx files can be found on RDSF:

```{r,}
# path to GTEx VCF file
gtex_vcf_dir <- ""
# path to expression & covariate tar directories downloaded from GTEx Portal 
covariate_matrix_tar  <- ""
expression_matrix_tar <- ""
```
alternitivly the GTEx directory on RDSF (should) contain a config.R file which can be sourced to attach these path variables:

```{r,}
source('config.R')
```
"covariate_matrix_tar" and "expression_matrix_tar" are matricies that contain normalised, QC'd data and are downloadable from the GTEx Portal here:

http://www.gtexportal.org/static/datasets/gtex_analysis_v6/single_tissue_eqtl_data/GTEx_Analysis_V6_eQTLInputFiles_geneLevelNormalizedExpressionMatrices.tar.gz

http://www.gtexportal.org/static/datasets/gtex_analysis_v6/single_tissue_eqtl_data/GTEx_Analysis_V6_eQTLInputFiles_covariates.tar.gz


### Running the pipeline

In order to generate a polygenic risk score, a "query data.frame" containg information for the sentinel SNPs are required.

SNP|effect.allele|other.allele|beta
---|---|---|---
rs123|A|C|0.136
rs456|G|A|-0.324

The "TwoSampleMR" [https://mrcieu.github.io/TwoSampleMR/] R package can be used to generate this dataframe for a given outcome and a p-value threshold.   

Extract 'query' SNPs from GTEx VCF files:

```{r,}
gtex_query <- extract_query_snps_gtex(query_data_frame = df, gtex_vcf_dir = gtex_vcf_dir)
```

Calculate the polygenic risk score:

```{r,}
geno <- calculate_prs_gtex(query_data_frame = df, gtex_data_frame = gtex_query)
```

Expression and covariate matricies have been downloaded from the GTEx portal and saved as an R object on RDSF. This object contains fully normalised and filtered expression values for each tissue, using the same protochol used by the GTEx consortium. The covariate matrix include the following; array, sex, 3 PCs and 15 PEER componants. See Aguet et al 2016 (https://doi.org/10.1101/074450) for more information on these data.

Load covariate and expression objects:

```{r,}

```

Make a character vector of the tissues you want to analyse:

```{r,}

```

RuneQTL analysis

Leave the "tx" argument out if you want to analyse all transcripts. Likewise, ignore the tissue argument if ypou wish to analyse all tissues: 

```{r,}
results <- run_eqtl(geno = geno, expression = exp_cov_list, tissue = tissue, tx = 1:500)
```
Render volcanoe plots with gene symbols added to significant genes:

```{r, annotation}
vp <- volcanoeplot(output, fdr = 0.05, outdir = '../../output/BMI/')
```
